---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# templateICAr

<!-- badges: start -->
[![R-CMD-check](https://github.com/mandymejia/templateICAr/workflows/R-CMD-check/badge.svg)](https://github.com/mandymejia/templateICAr/actions)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/mandymejia/templateICAr?branch=master&svg=true)](https://ci.appveyor.com/project/mandymejia/templateICAr)
[![Codecov test coverage](https://codecov.io/gh/mandymejia/templateICAr/branch/master/graph/badge.svg)](https://app.codecov.io/gh/mandymejia/templateICAr?branch=master)
<!-- badges: end -->

This package contains functions implementing the template ICA model proposed in Mejia et al. (2019) and the spatial template ICA model proposed in proposed in Mejia et al. (2020+). For both models, subject-level brain networks are estimated as deviations from known population-level networks, which can be estimated using standard ICA algorithms. Both models employ an expectation-maximization algorithm for estimation of the latent brain networks and unknown model parameters.

Template ICA consists of three steps. The main functions associated with each step are listed below. 

1. Template estimation: `estimate_template`. Can export the results with `export_template`.
2. Template ICA model estimation (single-subject): `templateICA`.
3. Identification of areas of engagement in each IC (or deviation from the template mean): `activations`.

## Citation

If you use `templateICAr`, please cite the Template ICA [paper](https://doi.org/10.1080/01621459.2019.1679638) as well as the spatial model [paper](https://doi.org/10.1080/10618600.2022.2104289) if applicable. You can also obtain citation information from within R like so:

``` {r}
citation("templateICAr")
```

## Installation

You can install the development version of `templateICAr` from Github with:

``` r
# install.packages("devtools")
devtools::install_github("templateICAr")
```


## Important Notes on Dependencies:

To analyze or visualize CIFTI-format data, `templateICAr` depends on the `ciftiTools` package, which requires an installation of Connectome Workbench.  It can be installed from the [HCP website](https://www.humanconnectome.org/software/get-connectome-workbench).

For fitting *spatial* template ICA model, INLA is required, along with an INLA-PARDISO license.  INLA is NOT required for running standard template ICA.  Due to a CRAN policy, INLA cannot be installed automatically. You can obtain it by running `install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)`.  Binaries for alternative Linux builds can be added with the command `inla.binary.install()`.  To obtain an INLA-PARDISO license, run `inla.pardiso()` in R after running `library(INLA)`. Once you obtain a license, point to it using `INLA::inla.setOption(pardiso.license = "pardiso.lic")` followed by `INLA::inla.pardiso.check()` to ensure that PARDISO is successfully installed and running.
